A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-04-11 ICLP Paper/samplepaper.tex
update a table
complete updating table 3 & 4
added sentence reacting to Reviewer 2

The \lstinline{step(w)} subprogram until line~\ref{prg:diff_log:oper_limit:end} 
in Listing~\ref{prg:encoding} constitutes the central part of our multi-shot ASP modulo DL encoding, whose parameter~\lstinline{w} stands for consecutive integers from~\lstinline{1}
identifying time windows to schedule.
Auxiliary atoms of the form \lstinline{use(}$m$\lstinline{,}$w'$\lstinline{,}$w$\lstinline{)},
supplied by the rules in lines~\ref{prg:encoding:use1} and~\ref{prg:encoding:use2},
indicate the latest time window $1\leq w'\leq w$ including some operation that needs
to be processed by machine~$m$.
The next rule in lines \ref{prg:base:sharing_machine:begin}-\ref{prg:base:sharing_machine:end}
identifies pairs \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} and
\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)} of operations sharing
the same machine~$m$,
where
\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)} belongs to the time window~$w$ and
\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} is either 
(i)
contained in the latest time window $1\leq w'< w$ indicated by
\lstinline{use(}$m$\lstinline{,}$w'$\lstinline{,}$w-1$\lstinline{)} or
(ii)
also part of the time window~$w$, in which case $j_1 < j_2$ establishes
an asymmetric representation for the pair of operations in derived atoms
\lstinline{share((}$j_1$\lstinline{,}$s_1$\lstinline{),(}$j_2$\lstinline{,}$s_2$\lstinline{),}$p_1$\lstinline{,}$p_2$\lstinline{,}$x$\lstinline{,}$w$\lstinline{)}.
If the flag $x=\text{\lstinline{1}}$ signals that 
\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)}
belongs to a previous time window~$w'$,
the rule in line~\ref{prg:base:ordered} % applies and 
derives the atom
\lstinline{order((}$j_1$\lstinline{,}$s_1$\lstinline{),(}$j_2$\lstinline{,}$s_2$\lstinline{),}$p_1$\lstinline{,}$w$\lstinline{)}
to signal that \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} needs to be completed
before performing \lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}, i.e.,
the execution order must comply with the decomposition into time windows.
The rule in lines \ref{prg:base:same_job:begin}-\ref{prg:base:same_job:end} yields
a similar atom when \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)}
is the predecessor operation $s_1 = s_2-1$ of \lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}
in the same job $j_1=j_2$.
In contrast to the cases in which 
\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} must be processed before
\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)},
the choice rule in line~\ref{prg:base:order}
allows for performing two operations sharing a machine in the lexicographic order
of their jobs if the operations belong to the same time window.
In case the atom representing execution in lexicographic order is not chosen,
the rule in lines \ref{prg:base:order:begin}-\ref{prg:base:order:end}
derives an atom signaling the inverse, given that the operations must not intersect and some sequence has to be determined.%
%
\lstinputlisting[float=t,label=prg:encoding,caption={Multi-shot ASP modulo DL encoding}]{listing/encoding.lp}

@@

    30\%             & $3051.2$  & $1000.0$   & $3.0$	& $3451.8$      & $1000.0$ & $4.0$ & $6933.3$ & $1000.0$ & $6.0$ % \\ 
%                     [1.5mm]
%                     
%    \\[-2.75mm]
%    
%    \multicolumn{10}{c}{\textbf{Compressing Time Windows}}\\
%    \hline
%    \multirow{2}*{Overlap} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
%     & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
%     \hline\\[-2.75mm]
    \midline
    0\%              & $3013.3$  & $972.3$   & $3.7$	   & $3186.8$      & $940.8$ & $3.3$ & $5713.0$ & $989.2$ & $8.6$\\
    % [1.5mm]

    10\%             & $2970.2$  & $1000.0$   & $4.0$	   & $3149.8$      & $1000.0$ & $4.0$ & $5656.0$ & $1000.0$ & $9.0$\\

    20\%             & $2952.2$  & $1000.0$   & $4.0$	   & \boldmath{$3147.7$}      & $1000.0$ & $4.0$ & \boldmath{$5619.2$} & $1000.0$ & $9.0$\\

    30\%             & \boldmath{$2943.1$}  & $1000.0$   & $4.0$	   & $3162.2$      & $1000.0$ & $4.0$ & $5857.0$ & $1000.0$ & $9.0$%\\ 
                     % [1.5mm]
    \botline %\hline
    \end{tabular}
   % }
%    \end{minipage}
% }
%    \hfill\hfill\mbox{~}
\end{table}
%
Focusing again on the static M-EST decomposition strategy,
Table~\ref{tab:Table03} assesses different percentages of overlapping
operations from previous time windows as well as the impact of compressing
(optimized) partial schedules in a postprocessing phase.
The baseline approach as above is included in the first row with 0\% of the
operations per time window overlap and without the use of compression.
The next three rows exclude compression as well but introduce 10\%, 20\%, or
30\% of overlapping operations that can be rescheduled together with operations
of the next time window.
In view of the considered number of time windows, picked according to the shortest
average makespan achieved with 0\% overlap, between $18$ and $35$ operations constitute
an overlap of 10\%, so that 30\% range up to more than $100$ operations.
Such non-negligible amounts can make the optimization of subproblems for time windows
harder, while the moderate overlaps of 10\% or 20\% manage to consistently
improve the solution quality.
Even more significant advantages are obtained by applying compression in the last
four rows,
where the average makespan gets already shorter with 0\% overlap.
Further improvements of roughly 2\% in solution quality are achieved with 20\% overlap,
which yield consistently better schedules than the smaller overlap of 10\% and
avoid deteriorations as observed with 30\% overlap on the largest instances
with $100\times20$ jobs and machines.
In summary, a Machine-based problem 
decomposition strategy along with
postprocessing to compress partial schedules as well as a modest overlap of about
20\% of the time window size turn out to be most successful for the JSP
benchmark sets under consideration.

@@

\begin{table}[t]
  \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
%    \mbox{~}\hfill
%    \resizebox{0.8\textwidth}{!}{%
%    \begin{minipage}{\textwidth}
  \centering
  \begin{tabular}{l l r r }

@@

  \botline %\hline
  \end{tabular}
\end{table}
%

>>>>

The \lstinline{step(w)} subprogram until line~\ref{prg:diff_log:oper_limit:end} 
in Listing~\ref{prg:encoding} constitutes the central part of our multi-shot ASP modulo DL encoding, whose parameter~\lstinline{w} stands for consecutive integers from~\lstinline{1}
identifying time windows to schedule.
Auxiliary atoms of the form \lstinline{use(}$m$\lstinline{,}$w'$\lstinline{,}$w$\lstinline{)},
supplied by the rules in lines~\ref{prg:encoding:use1} and~\ref{prg:encoding:use2},
indicate the latest time window $1\leq w'\leq w$ including some operation that needs
to be processed by machine~$m$.
The next rule in lines \ref{prg:base:sharing_machine:begin}-\ref{prg:base:sharing_machine:end}
identifies pairs \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} and
\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)} of operations sharing
the same machine~$m$,
where
\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)} belongs to the time window~$w$ and
\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} is either 
(i)
contained in the latest time window $1\leq w'< w$ indicated by
\lstinline{use(}$m$\lstinline{,}$w'$\lstinline{,}$w-1$\lstinline{)} or
(ii)
also part of the time window~$w$, in which case $j_1 < j_2$ establishes
an asymmetric representation for the pair of operations in derived atoms
\lstinline{share((}$j_1$\lstinline{,}$s_1$\lstinline{),(}$j_2$\lstinline{,}$s_2$\lstinline{),}$p_1$\lstinline{,}$p_2$\lstinline{,}$x$\lstinline{,}$w$\lstinline{)}.
If the flag $x=\text{\lstinline{1}}$ signals that 
\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)}
belongs to a previous time window~$w'$,
the rule in line~\ref{prg:base:ordered} % applies and 
derives the atom
\lstinline{order((}$j_1$\lstinline{,}$s_1$\lstinline{),(}$j_2$\lstinline{,}$s_2$\lstinline{),}$p_1$\lstinline{,}$w$\lstinline{)}
to express that \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} needs to be completed
before performing \lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}, i.e.,
the execution order must comply with the decomposition into time windows.
The rule in lines \ref{prg:base:same_job:begin}-\ref{prg:base:same_job:end} yields
a similar atom when \lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)}
is the predecessor operation $s_1 = s_2-1$ of \lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}
in the same job $j_1=j_2$.
In contrast to the cases in which 
\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} must be processed before
\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)},
the choice rule in line~\ref{prg:base:order}
allows for performing two operations sharing a machine in the lexicographic order
of their jobs if the operations belong to the same time window.
In case the atom representing execution in lexicographic order is not chosen,
the rule in lines \ref{prg:base:order:begin}-\ref{prg:base:order:end}
derives an atom expressing the inverse, given that the operations must not intersect and some sequence has to be determined.
\hl{Note that applications of the rules in lines} \ref{prg:base:order}-\ref{prg:base:order:end}
\hl{for choosing an execution order require that the
operations}
\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} \hl{and}
\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}
\hl{are part of the same time window~$w$ and belong to distinct jobs, so that}
\lstinline{(}$j_1$\lstinline{,}$s_1$\lstinline{)} \hl{and}
\lstinline{(}$j_2$\lstinline{,}$s_2$\lstinline{)}
\hl{are out of scope of the rules
imposing a fixed order in lines} \ref{prg:base:ordered}-\ref{prg:base:same_job:end}.%
\lstinputlisting[float=t,label=prg:encoding,caption={Multi-shot ASP modulo DL encoding}]{listing/encoding.lp}

@@

    30\%             & $3051.2$  & $1000.0$   & $3.0$	& $3451.8$      & $1000.0$ & $4.0$ & $6933.3$ & $1000.0$ & $6.0$ % \\ 
%                     [1.5mm]
%                     
%    \\[-2.75mm]
%    
%    \multicolumn{10}{c}{\textbf{Compressing Time Windows}}\\
%    \hline
%    \multirow{2}*{Overlap} & \multicolumn{3}{c}{50$\times$15} & \multicolumn{3}{c}{50$\times$20} & \multicolumn{3}{c}{100$\times$20}\\
%     & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC & Makespan & CPU(Sec) & IC \\
%     \hline\\[-2.75mm]
    \midline
    0\%              & $3010.3$  & $907.5$   & $3.4$	   & $3174.5$      & $839.6$ & $3.0$ & $5671.9$ & $1000.0$ & $7.0$\\
    % [1.5mm]

    10\%             & $2960.6$  & $989.4$   & $3.9$	   & $3154.6$      & $956.4$ & $3.9$ & \boldmath{$5637.5$} & $1000.0$ & $7.0$\\

    20\%             & $2946.9$  & $1000.0$   & $4.0$	   & \boldmath{$3120.3$}      & $1000.0$ & $4.0$ & $5640.7$ & $1000.0$ & $7.0$\\

    30\%             & \boldmath{$2932.9$}  & $1000.0$   & $4.0$	   & $3169.0$      & $1000.0$ & $4.0$ & $5722.7$ & $1000.0$ & $7.0$%\\ 
                     % [1.5mm]
    \botline %\hline
    \end{tabular}
   % }
%    \end{minipage}
% }
%    \hfill\hfill\mbox{~}
\end{table}
%
Focusing again on the static M-EST decomposition strategy,
Table~\ref{tab:Table03} assesses different percentages of overlapping
operations from previous time windows as well as the impact of compressing
(optimized) partial schedules in a postprocessing phase.
The baseline approach as above is included in the first row with 0\% of the
operations per time window overlap and without the use of compression.
The next three rows exclude compression as well but introduce 10\%, 20\%, or
30\% of overlapping operations that can be rescheduled together with operations
of the next time window.
In view of the considered number of time windows, picked according to the shortest
average makespan achieved with 0\% overlap, between $18$ and $35$ operations constitute
an overlap of 10\%, so that 30\% range up to more than $100$ operations.
Such non-negligible amounts can make the optimization of subproblems for time windows
harder, while the moderate overlaps of 10\% or 20\% manage to consistently
improve the solution quality.
Even more significant advantages are obtained by applying compression in the last
four rows,
where the average makespan gets already shorter with 0\% overlap.
Further improvements of roughly 2\% in solution quality are achieved with 20\% overlap,
which yield consistently better schedules than the smaller overlap of 10\% and
avoid deteriorations as observed with 30\% overlap on the largest instances
with $100\times20$ jobs and machines.
In summary, a Machine-based problem 
decomposition strategy along with
postprocessing to compress partial schedules as well as a modest overlap of about
20\% of the time window size turn out to be most successful for the JSP
benchmark sets under consideration.

@@

\begin{comment}
  \begin{table}[t]
  \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
  \centering
  \begin{tabular}{l l r r }

@@

  \botline %\hline
  \end{tabular}
\end{table}
%
\end{comment}

\begin{table}[t]
  \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
  \centering
  \begin{tabular}{l l r r r r r r}

  \topline%\hline
  \multirow{3}{*}{Instances} & \multicolumn{4}{c}{Makespan} & \multicolumn{3}{c}{Distance measures} \\
  & Optima & OR-tools & \multicolumn{2}{c}{ASP} & OR-tools & \multicolumn{2}{c}{ASP} \\

  & & & Single-shot & Multi-shot &  & Single-shot & Multi-shot \\

  $50\times15$ & $2733.8$ & $2778.8$ & $3542.1$ & $2932.9$ & $5.0$ & $768.3$ & $159.1$ \\
  $50\times20$ & $2843.9$ & $2914.3$ & $3879.4$ & $3120.3$ & $70.4$ & $1035.5$ & $276.4$ \\
  $100\times20$ & $5365.7$ & $5559.3$ & $46786.1$ & $5637.5$ & $193.6$ & $41420.4$ & $271.8$

  \botline %\hline
  \end{tabular}
\end{table}
